
---
title: "Markdown Template of post"
output:
  md_document:
    variant: markdown_github+tex_math_dollars+autolink_bare_uris+ascii_identifiers
    pandoc_args: [
      "--latexmathml"
    ]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
# plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Including Formulas

$$ \sum_{i = 1}^n \frac{1}{i} = \infty $$

Obviously since $\int_{\mathbb{R}} \frac{1}{x} dx = + \infty$.

## Including code

    if (a > 3) {
      moveShip(5 * gravity, DOWN);
    }


