---
layout: post
title: "Applying diffusion theory to Google Trends"
subtitle: on example of Candy Crush Saga adoption
bigimg: /img/math.jpg
output:
  html_document:
    toc: no
  md_document:
    pandoc_args: --latexmathml
    variant: markdown_github+tex_math_dollars+autolink_bare_uris+ascii_identifiers
tags: R
---

```{r setup, include=FALSE}

  # library(knitr)
  # options(markdown.HTML.options = NULL)
  # 
  # # knitr options
  # knitr::opts_chunk$set(echo = TRUE)
  # post.title <- "markdown-template"
  # post.date <- Sys.Date()
  # fig.path <- paste0("{{ site.url }}/img/", post.title, "-")
  # knitr::opts_chunk$set(fig.path=fig.path)
  
```

## Latex inline (must be in single line)

where $b = 1$

where $\sum_{i = 1}^n \frac{1}{i} = \infty$ and

where $\int_0^\infty \mathrm{e}^{-x}\,\mathrm{d}x$ and

where $S_t$ and

where $s_t$ and

where $e = e^{-\eta}$ and 

where $\sqrt\pi$ and 

where $\sqrt{\frac{1}{2}}$ and 

## Latex equations

This is an R Markdown format used for publishing markdo
$$ b = 1 $$
This is an R Markdown format used for publishing markdo
$$ b = \frac{1}{2} $$
This is an R Markdown format used for publishing markdo
$$ \int_0^\infty \mathrm{e}^{-x}\,\mathrm{d}x $$
This is an R Markdown format used for publishing markdo
$$ (X, U) \sim \mathbb{U}\{(x, u) : 0 < u < f(x)\} $$



## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Including Formulas

$$ = \frac{1}{M(b, \eta)} \left\{ 1_{[0, t^*]} * m(b, \eta) + 1_{(t^*, +\infty)} * (1 + \eta) * b \exp(-bt) \right\} $$

$$ = \frac{m(b,\eta) t^*}{M(b,\eta)} * \frac{1}{t^*}1_{[0, t^*]}  + \frac{m(b,\eta) e^{-bt^*}(1+\eta)}{M(b,\eta)} * 1_{(t^*, +\infty)} b \exp(-b(t - t^*))$$

$$ = p_1(b,\eta) * \mathbb{U}[0, t^*] + p_2(b, \eta) * \mathrm{Exp}(b)|_{t^*}^{\infty} $$


Obviously since $\int_{\mathbb{R}} \frac{1}{x} dx = + \infty$ as in [^1]

## Including code


    if (a > 3) {
      moveShip(5 * gravity, DOWN);
    }


[^1]: Abramovitz and Steingun
